A Formal Model of Checked C

### 2022-06-25 sandp23/background.tex
update.

\documentclass[conference]{IEEEtran}

\makeatletter
\def\endthebibliography{\def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}\endlist
}
\makeatother

\IEEEpeerreviewmaketitle

\usepackage{filecontents}
\usepackage[noadjust]{cite}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{ebproof}
\usepackage{todonotes}
\usepackage{amsfonts}
\usepackage{algorithmic}
\usepackage{textcomp}
\usepackage{mathpartir}
\usepackage{color}
\usepackage{xcolor}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\usepackage{subcaption} \usepackage{epsfig}
\usepackage{listings}
\usepackage[override]{cmtt}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{url}
\usepackage{xparse}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{chains,fit,shapes,positioning,calc,arrows.meta,shapes.arrows}

\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}

\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  numbers=left,
  xleftmargin=\parindent,
  language=C,
  columns=flexible,      
  showstringspaces=false,
  basicstyle=\small\sffamily,
  otherkeywords={uint},
  literate={{<-}{{$\leftarrow\,$}}2
            {->}{{$\rightarrow\,$}}2
            {<=}{{$\leq\,$}}2},
  numberstyle=\tiny\rmfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue!80!black},
}
\lstset{language=C,style=customc}

\newenvironment{DIFnomarkup}{}{}
\newcommand{\myparagraph}[1]{\textbf{#1}.\xspace}

\newcommand{\code}[1]{\lstinline|#1|}
%\newcommand{\ignore}[1]{{}}
\newcommand{\kwchecked}{\lstinline|\_Checked|}
\newcommand{\kwunchecked}{\lstinline|\_Unchecked|}
\newcommand{\kwdynamiccheck}{\lstinline|\_Dynamic\_check|}

\newcommand{\var}[1]{#1}
\newcommand{\Arrayptr}[1]{\lstinline|_Array_ptr<|{#1}\lstinline|>|}
\newcommand{\ArrayptrT}{\Arrayptr{$\tau$}}
\newcommand{\Ptr}[1]{\lstinline|_Ptr<|{#1}\lstinline|>|}
\newcommand{\PtrName}{\lstinline|_Ptr|}
\newcommand{\PtrT}{\Ptr{$\tau$}}
\newcommand{\Ntarrayptr}[1]{\lstinline|_Nt_array_ptr<|{#1}\lstinline|>|}
\newcommand{\NtarrayptrT}{\Ntarrayptr{$\tau$}}
\newcommand{\Checkedarr}[2]{{#1}~\lstinline|_Checked[|{#2}\lstinline|]|}
\newcommand{\CheckedarrT}[1]{\Checkedarr{$\tau$}{#1}}

\newif\ifsubmit\submitfalse

\ifsubmit
\newcommand{\mwh}[1]{}
\newcommand{\ashe}[1]{}
\newcommand{\dtarditi}[1]{}
\newcommand{\dvh}[1]{}
\newcommand{\leo}[1]{}
\newcommand{\liyi}[1]{}
\newcommand{\yiyun}[1]{}
\else
\newcommand{\mwh}[1]{\textcolor{red}{Mike: #1}}
\newcommand{\dtarditi}[1]{\textcolor{purple}{David: #1}}
\newcommand{\leo}[1]{\textcolor{green!80!blue}{Leo: #1}}
\newcommand{\liyi}[1]{\textbf{\textcolor{orange}{[ Liyi: #1 ]}}}
\newcommand{\yiyun}[1]{\textcolor{cyan}{Yiyun: #1}}
\newcommand{\dvh}[1]{\textcolor{magenta}{DVH: #1}}
\fi

\newcommand{\lang}{\textsc{CoreChkC}\xspace}
\newcommand{\elang}{\textsc{CoreC}\xspace}

\newtheorem{defi}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{prop}{Proposition}

\newcommand{\CoreChkC}{\textsc{CoreChkC}\xspace}
\newcommand{\checkedc}{\text{Checked C}\xspace}

 \newcommand{\rulelab}[1]{{\small \textsc{#1}}}
\newcommand{\steps}{\ensuremath{\longrightarrow}}
\newcommand{\tbool}{\texttt{bool}}
\newcommand{\tsizeof}{\texttt{sizeof}}
\newcommand{\vbool}{\texttt{Bool}}
\newcommand{\verror}{\texttt{Error}}
\newcommand{\vval}{\mathpzc{val}}
\newcommand{\tfixed}{\texttt{fixedp}}
\newcommand{\tnat}{\texttt{nat}}
\newcommand{\tarr}[2]{\texttt{array}~{#1}~{#2}}
\newcommand{\econst}[2]{({#1}){#2}}
\newcommand{\eindex}[2]{{#1}\texttt{[}{#2}\texttt{]}}
\newcommand{\sassign}[4]{{#1} \leftarrow {#3}~{#2}~{#4}}
\newcommand{\ssassign}[3]{{#1} \xleftarrow{#2} {#3}}
\newcommand{\sif}[3]{\texttt{if}~{#1}~{#2}~{#3}}
\newcommand{\sfor}[3]{\texttt{for}~{#1}~{#2}~{#3}}
\newcommand{\scall}[3]{{#1}\leftarrow {#2}~{#3}}
\newcommand{\sseq}[2]{{#1}\,\texttt{;}\,{#2}}
\newcommand{\sinv}[1]{\texttt{inv}~{#1}}
\newcommand{\inst}[3][ ]{\texttt{#2}^{#1}~{#3}}
\newcommand{\insttwo}[4][ ]{\texttt{#2}^{#1}~{#3}~{#4}}
\newcommand{\instthree}[5][ ]{\texttt{#2}^{#1}~{#3}~{#4}~{#5}}
\newcommand{\iskip}[1]{\inst{ID}{#1}}
\newcommand{\inot}[1]{\inst{X}{#1}}
\newcommand{\ictrl}[2]{\insttwo{CU}{#1}{#2}}
\newcommand{\irz}[3][ ]{\insttwo[#1]{RZ}{#2}{#3}}
\newcommand{\isr}[3][ ]{\insttwo[#1]{SR}{#2}{#3}}
\newcommand{\icnot}[2]{\insttwo{CNOT}{#1}{#2}}
\newcommand{\ilshift}[1]{\inst{Lshift}{#1}}
\newcommand{\irshift}[1]{\inst{Rshift}{#1}}
\newcommand{\irev}[1]{\inst{Rev}{#1}}
\newcommand{\iqft}[2][ ]{\inst[#1]{QFT}{#2}}
\newcommand{\ihad}[1]{\inst{H}{#1}}

\newcommand{\iseq}[2]{{#1}\,\texttt{;}\,{#2}}
\newcommand{\inval}[2]{\insttwo{Nval}{#1}{#2}}
\newcommand{\ihval}[3]{\instthree{Hval}{#1}{#2}{#3}}
\newcommand{\iqval}[2]{\insttwo{Qval}{#1}{#2}}

\newcommand{\instr}{\iota}

\newcommand{\kw}[1]{\ensuremath{\mathtt{#1}}}
\newcommand{\subtype}[2]{\ensuremath{\vdash{#1}\prec{#2}}}
\newcommand{\estrlen}[1]{\ensuremath{\kw{strlen}({#1})}}
\newcommand{\estrlentext}{\ensuremath{\kw{strlen}}}

\newcommand{\tarray}[3]{\tarrayb{({#1},{#2})}{#3}}
\newcommand{\tarrayb}[2]{\ensuremath{[{#1}~{#2}]}}
\newcommand{\tntarray}[3]{\tntarrayb{({#1},{#2})}{#3}}
\newcommand{\tntarrayb}[2]{\tarrayb{#1}{#2}_{nt}}
\newcommand{\tallarrayb}[2]{\ensuremath{[{#1}~{#2}]_{\kappa}}}
\newcommand{\tallarraybc}[2]{\ensuremath{[{#1}~{#2}]_{\textcolor{cyan}{\kappa}}}}

\newcommand{\tallarray}[3]{\tallarrayb{({#1},{#2})}{#3}}

\newcommand{\tptr}[2]{\ensuremath{\mathtt{ptr}^{#2}~{#1}}}
\newcommand{\tntptr}[4]{\ensuremath{[({#1},{#2})~{#3}]^{#4}_{nt}*}}

\newcommand{\tarrayptr}[4]{{\tptr{\tarray{#1}{#2}{#3}}{#4}}}
\newcommand{\tntarrayptr}[4]{{\tptr{\tntarray{#1}{#2}{#3}}{#4}}}
\newcommand{\tallarrayptr}[4]{{\tptr{\tallarray{#1}{#2}{#3}}{#4}}}

\newcommand{\tgarray}[4]{\ensuremath{\mathtt{#1}~{#2}~{#3}~{#4}}}
\newcommand{\tstruct}[1]{\ensuremath{\kw{struct}~{#1}}}

\newcommand{\evalue}[2]{\ensuremath{{#1}\!:\!{#2}}}

\newcommand{\emalloc}[1]{\ensuremath{\kw{malloc}({#1})}}
\newcommand{\emalloctext}{\ensuremath{\kw{malloc}}}
\newcommand{\ecall}[2]{\ensuremath{{#1}({#2})}}
\newcommand{\ret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
\newcommand{\rettext}{\ensuremath{\kw{ret}}}
\newcommand{\ecast}[2]{\ensuremath{\kw{(}{#1}\kw{)}{#2}}}
\newcommand{\edyncast}[2]{\ensuremath{\langle{#1}\rangle{#2}}}
\newcommand{\edcast}[2]{\ensuremath{\kw{(D},{#1}\kw{)}{#2}}}
\newcommand{\elet}[3]{\ensuremath{\kw{let}~#1\, \texttt{=}\, #2~\kw{in}\;{#3}}}
\newcommand{\elettext}{\ensuremath{\kw{let}}}
\newcommand{\ebinop}[2]{\ensuremath{#1 \plus #2}}
\newcommand{\eassign}[2]{\ensuremath{\texttt{*}{#1}\,\texttt{=}\, {#2}}}
\newcommand{\eassignstack}[2]{\ensuremath{{#1}\,\texttt{=}\, {#2}}}
\newcommand{\efield}[2]{\ensuremath{{#1}\kw{\rightarrow}{#2}}}
\newcommand{\estar}[1]{\ensuremath{\texttt{*}{#1}}}
\newcommand{\eamper}[2]{\ensuremath{\kw{\&}{#1}\kw{\rightarrow}{#2}}}
\newcommand{\eunchecked}[1]{\ensuremath{\kw{unchecked}\;{#1}}}
\newcommand{\euncheckedtext}{\ensuremath{\kw{unchecked}}}
\newcommand{\eif}[3]{\ensuremath{\kw{if\;}(#1)\;{#2}\;\kw{else}\;{#3}}}
\newcommand{\eifa}[3]{\ensuremath{\kw{IF\;}(#1)\;{#2}\;\kw{ELSE}\;{#3}}}
\newcommand{\eifatext}{\ensuremath{\kw{IF}}}

\newcommand{\eiftext}{\ensuremath{\kw{if}}}
\newcommand{\ethentext}{\ensuremath{\kw{then}}}
\newcommand{\eelse}{\ensuremath{\kw{else}}}
\newcommand{\ebreak}{\ensuremath{\kw{break}}}
\newcommand{\ebounds}{\ensuremath{\kw{bounds}}}
\newcommand{\enull}{\ensuremath{\kw{null}}}
\newcommand{\hole}{\ensuremath{\Box}}
\newcommand{\defscope}{\sigma}
\newcommand{\wf}[1]{wf \, #1}
\newcommand{\tint}{\ensuremath{\mathtt{int}}}
\newcommand{\efor}{\ensuremath{\mathtt{for}}}
\newcommand{\ewhile}{\ensuremath{\mathtt{while}}}
\newcommand{\etrue}{\ensuremath{\mathtt{true}}}
\newcommand{\efalse}{\ensuremath{\mathtt{false}}}
\newcommand{\edplus}{\ensuremath{\texttt{++}}}
\newcommand{\edminus}{\ensuremath{\texttt{--}}}

\newcommand{\ememcpy}{\ensuremath{\mathtt{memcpy}}}
\newcommand{\ereturn}{\ensuremath{\mathtt{return}}}
\newcommand{\emain}{\ensuremath{\mathtt{main}}}
\newcommand{\estrcat}{\ensuremath{\mathtt{strcat}}}
\newcommand{\estrcatbad}{\ensuremath{\mathtt{strcat\_b}}}
\newcommand{\heap}{\ensuremath{\mathpzc{H}}}
\newcommand{\eret}[3]{\ensuremath{\kw{ret}({#1},{#2},{#3})}}
\newcommand{\erettext}{\ensuremath{\kw{ret}}}

\newcommand{\plus}{\mathbin{\texttt{+}}}

\newcommand{\fv}{\mathit{FV}}
\newcommand{\dom}{\mathit{dom}}

\newcommand{\fm}{\mathit{FM}}
\newcommand{\size}{\mathit{size}}

\newcommand{\cmode}{\texttt{c}}
\newcommand{\umode}{\texttt{u}}
\newcommand{\bvar}{\ensuremath{\beta}}
\newcommand{\mode}{\textit{mode}}

\newcommand{\cextend}[4]{\ensuremath{#4 = ~ \vdash_{extend} #1, #2, #3}}
\newcommand{\fresh}{\ensuremath{\kw{fresh}}}
\newcommand{\echecknull}[3]{\ensuremath{#3 = ~\vdash_{null}#1, #2}}
\newcommand{\echeckboundsdyn}[5]{\ensuremath{#5 = ~ \vdash_{boundsD}#1, #2, #3, #4}}
\newcommand{\echeckbounds}[5]{\ensuremath{#5 = ~ \vdash_{boundsR}#1, #2, #3, #4}}
\newcommand{\echeckboundsw}[5]{\ensuremath{#5 = ~ \vdash_{boundsW}#1, #2, #3, #4}}
\newcommand{\esizeof}[1]{\ensuremath{\kw{sizeof}(#1)}}
\newcommand{\ewidenstrlen}[5]{\ensuremath{#5 = ~ \vdash_{widenstr}#1, #2, #3, #4}}
\newcommand{\ewidenderef}[4]{\ensuremath{#4 = ~ \vdash_{widenderef}#1, #2, #3}}
\newcommand{\eleq}[2]{\ensuremath{#1 \leq #2}}

\newcommand{\rboundle}[2]{\ensuremath{\vdash_{bounds}#1 \leq #2}}

\def\codesize{\normalsize}
\definecolor{programs}{gray}{0.1}

\lstset{
  basicstyle=\ttfamily,  
}

\makeatletter
\lst@CCPutMacro
    \lst@ProcessOther {"22}{\lst@ifupquote \textquotedbl
                                     \else \char34\relax \fi}
    \@empty\z@\@empty
\makeatother

\lstloadlanguages{C}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{A Formal Model of Checked C}

% \ignore{
% \author{\IEEEauthorblockN{1\textsuperscript{st} Leonidas Lampropoulos}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Maryland}\\
% llampro@cs.umd.edu}
% \and
% \IEEEauthorblockN{1\textsuperscript{st} Michael Hicks}
% \IEEEauthorblockA{\textit{Department of Computer Science} \\
% \textit{University of Maryland}\\
% llampro@cs.umd.edu}
% }
% \author{Leonidas Lampropoulos \and
% Michael Hicks}
% }
\maketitle

\begin{abstract}

  We present a formal model of Checked C, a dialect of C
  that aims to enforce spatial memory safety. Our model pays particular
  attention to the semantics of dynamically sized, potentially null-terminated
  arrays.
We formalize this model in Coq, and prove that any spatial memory
  safety errors can be \emph{blamed} on portions of the program
  labeled \emph{unchecked}; this is a Checked C feature that supports
  incremental porting and backward compatibility.
Our model develops an operational semantics that uses fat pointers to
  guarantee spatial safety. However, we formalize a compilation scheme
  that can yield thin pointers, with bounds information managed using
  inserted code. We show that the generated code faithfully simulates
  the original.
Finally, we build an executable version of our model in PLT Redex, 
  and use a custom random generator
  for well-typed and almost-well-typed terms to find
  inconsistencies between our model and the Clang Checked C
  implementation. We find these steps to be a useful way to co-develop a language
  (Checked C is still in development) and a core model of it.
\end{abstract}

\section{Introduction}\label{sec:intros}

The C programming language remains extremely popular despite the
emergence of new, modern languages. Unfortunately, C programs lack
spatial memory safety, which makes them susceptible to a host
of devastating vulnerabilities, including buffer overflows and
out-of-bounds reads/writes. Despite their long history, buffer
overflows and other spatial safety violations are among the most
prevalent and dangerous vulnerabilities on the Internet today \cite{Zeng:2013:SRF:2534766.2534798}.

Several industrial and research efforts---including CCured~\cite{Necula2005},
Softbound~\cite{softbound}, and ASAN~\cite{Serebryany2012}---have
explored means to compile C programs 
to automatically enforce spatial safety. These
approaches all impose performance overheads that are deemed too high for
use in deployment. Recently, Microsoft introduced \checkedc, an
open-source extension to C with new types and
annotations whose use can ensure a program’s spatial
safety~\cite{Elliott2018}. Importantly, \checkedc supports development that is
incremental and compositional. Code regions (e.g.,
functions or whole files) designated as \emph{checked} are sure to enforce
spatial safety, a property which is preserved via composition with
other checked regions. But not all regions must be checked: Checked
C's annotated \emph{checked pointers} are binary-compatible with legacy pointers, and
may coexist in the same code, which permits a deliberate (and
semi-automated) refactoring process. Parts of the FreeBSD kernel have
been successfully ported to \checkedc~\cite{duanrefactoring}, and overall, performance
overhead seems low enough for practical deployment.

While \checkedc promises to enforce spatial safety, we might wonder
whether its design and implementation deliver on this promise, or even
what ``spatial safety'' means when a program contains both checked and
unchecked code. In prior work, \citet{ruef18checkedc-incr} developed a core
formalization of \checkedc and with it proved that \emph{checked code cannot 
  be blamed}: any spatial safety violation can only be attributed to
code that is not in a checked region. While their work is a good start,
it fails to model important aspects of \checkedc's functionality,
particularly those involving pointers to arrays. In this paper, we
cover this gap, making three main contributions.

\myparagraph{Dynamically bounded and null-terminated arrays}
Our first contribution is a core formalism called \lang, which extends
\citet{ruef18checkedc-incr} with several new features, most notably 
\emph{dynamically bounded arrays} (Section~\ref{sec:formal}). 
Dynamically bounded arrays are those whose size is
known only at run time, as designated by in-scope variables using
dependent types. A pointer's accessible memory is bounded both above
and below, to admit arbitrary pointer arithmetic.

\lang also models
\emph{null-terminated} arrays, a kind of dynamically bounded array whose upper bound 
defines the array's \emph{minimum} length---additional space is available 
up to a null terminator. For example, the \checkedc type
\lstinline{nt_array_ptr<char> p:count(n)} says that
\lstinline{p} has length \emph{at least} \lstinline{n} (excluding the
null terminator), but further capacity is present if \lstinline|p[n]| is not null. \checkedc (and \lang) supports flow-sensitive
\emph{bounds widening}: statements of the form \lstinline|if (*p) | $s$,
where \lstinline{p}'s type is \lstinline{nt_array_ptr<T> count(0)},
typecheck statement $s$ under the assumption that \lstinline{p} has
type \lstinline{nt_array_ptr<T> count(1)}, 
 i.e., one more than it was,
since the character at the current-known length is
non-null. Similarly, the call \lstinline|n = strlen(p)| will widen
\lstinline|p|'s bounds to \lstinline|n|. Subtyping permits treating
null-terminated arrays as normal arrays of the same size (which
does not include, and thereby protects, the null terminator).

We prove, in Coq, a blame theorem for \lang.  As far as we are aware,
ours is the first formalized type system and proof of soundness for
pointers to null-terminated arrays with expandable bounds.

\myparagraph{Sound compilation of checked pointers} Our second
contribution is a formalization of bounds-check insertion for array
accesses (Section~\ref{sec:compilation}). Our operational semantics
annotates each pointer with metadata that describes its bounds, and
the assignment and dereference rules have premises to confirm the
access is in bounds. An obvious compilation scheme (taken by
Cyclone~\cite{Jim2002,GrossmanMJHWC02}, CCured~\cite{Necula2005}, and
earlier works) would be to translate annotated pointers to multi-word
objects: one word for the pointer, and 1-2 words to describe its lower
and upper bounds. Inserted checks reference these bounds. While
convenient, such ``fat'' pointers are expensive, and break backward
binary compatibility with legacy pointers. We formalize \checkedc's
compilation approach, which uses a single machine word for the
pointer, and adds checks involving the declared bounds (e.g., in a
dependent type) or additional stack-allocated \emph{ghost variables}
to accommodate bounds widening. We show that the compiled program
\emph{simulates} the original by mechanizing \lang and the compilation
judgment in PLT Redex~\cite{pltredex}, and use its random testing
feature to give confidence that simulation holds.

As far as we are aware, \lang is the first formalism to cleanly
separate bounds-checking compilation from the core semantics; prior
work merged the two, conflating \emph{meaning} with
\emph{mechanism}~\cite{Feng2006,Condit2007}. In carrying out the 
formalization, we discovered that our compilation approach is 
more expressive than that proposed in the Checked C
specification~\cite{checkedc} (Section~\ref{sec:disc}); we 
would not have discovered this improvement had we not separated checks from 
semantics.

\myparagraph{Model-based randomized testing} Our third
and final contribution is a strategy and implementation of model-based
randomized testing (Section~\ref{sec:evaluation}).  To check the
correctness of our formal model, we compare the behavior between the
existing Clang \checkedc implementation and our own model. This is
done by a conversion tool that converts expressions from \lang into
actual \checkedc code that can be compiled by the Clang \checkedc
compiler. We build a random generator of programs largely based on the
typing rules of \lang and make sure that, both statically and
dynamically, \lang and Clang Checked C are consistent after
conversion.  This helped rapidly prototype the model and uncovered
several issues in the Checked C compiler.

We begin with a review of \checkedc (Section~\ref{sec:overview}),
present our main contributions (Sections~\ref{sec:formal}--\ref{sec:evaluation}),
and conclude with a discussion of
related and future work (Sections~\ref{sec:related}, \ref{sec:conclude}).

\section{\checkedc Overview}\label{sec:overview}

This section describes Checked C, which extends C with new pointer
types and annotations that ensure spatial safety. More
details can be found in a prior overview~\cite{Elliott2018} or the
full specification~\cite{checkedc}.
Checked C is implemented as a fork of Clang/LLVM and is freely
available.\footnote{\url{https://github.com/Microsoft/checkedc-clang}}

\begin{figure}[t]
{\small
  \begin{lstlisting}[xleftmargin=4 mm]
nt_array_ptr<const char>
parse_utf16_hex(nt_array_ptr<const char> s,
                  ptr<uint> result) {
 int x1, x2, x3, x4;
 if (s[0] != 0) { x1 = hex_char_to_int(s[0]);
 if (s[1] != 0) { x2 = hex_char_to_int(s[1]);
 if (s[2] != 0) { x3 = hex_char_to_int(s[2]);
 if (s[3] != 0) { x4 = hex_char_to_int(s[3]);
 if (x1 != -1 && x2 != -1 && x3 != -1 && x4 != -1){
   *result = (uint)((x1<<12)|(x2<<8)|(x3<<4)|x4);
   return s+4;
  ...// several } braces
 } 
 return 0;
}
void parse(nt_array_ptr<const char> s,
            array_ptr<uint> p : count(n), 
            int n) {
 array_ptr<uint> q : bounds(p,p+n) = p;
 while (s && q < p+n) {
   array_ptr<uint> r : count(1) =
     dyn_bounds_cast<array_ptr<uint>>(q,count(1));
   s = parse_utf16_hex(s,r);
   q++;
 }
}
  \end{lstlisting}

}

\caption{Parsing a string of UTF16 hex characters in \checkedc}
\label{fig:checkedc-example}
\end{figure}

\subsection{Checked Pointer Types}
\checkedc introduces three varieties of \emph{checked pointer}:
\begin{itemize}
\item \code{ptr<}$T$\code{>} types a pointer that is either null or
  points to a single object of type $T$.
\item \code{array_ptr<}$T$\code{>} types a pointer that is either null
  or points to an array of $T$ objects. The array width is defined
  by a \emph{bounds} expression, discussed below.
\item \code{nt_array_ptr<}$T$\code{>} is like
  \code{array_ptr<}$T$\code{>} except that the bounds expression
  defines the \emph{minimum} array width---additional objects may
  be available past the upper bound, up to a null terminator.
\end{itemize}
A bounds expression used with the latter two pointer types has three
forms:
\begin{itemize}
\item \code{count(}$e$\code{)} where $e$ defines the array's
  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
  expression $e$ must be side-effect free and may only refer to
  variables whose addresses are not taken, or adjacent \code{struct}
  fields.
\item \code{byte_count(}$e$\code{)} is like \code{count}, but
  expresses arithmetic using bytes, no objects; i.e.,
  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
  pointers that bound the accessible region $[e_l,e_h)$ (the
  expressions are similarly restricted). Bounds
  \code{count(}$e$\code{)} is shorthand for
  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
  expression is useful for supporting pointer arithmetic.
\end{itemize}
The \checkedc compiler will instrument loads and stores of checked
pointers to confirm the pointer is non-null, and the access is within
the specified bounds. For pointers $p$ of type
\code{nt_array_ptr<}$T$\code{>}, such a check could spuriously fail if
the index is past $p$'s specified upper bound, but before the null
terminator. To address this problem, Checked C supports \emph{bounds
  widening}. If $p$ has bounds expression
\code{bounds(}$e_l$,$e_h$\code{)} a program may read from (but not
write to) $e_h$; when the compiler notices that a non-null character
is read at the upper bound, it will extend that bound to $e_h+1$.

\subsection{Example}

Fig.~\ref{fig:checkedc-example} gives an example \checkedc
program.\footnote{Ported from the Parson JSON
parser, \url{https://github.com/kgabis/parson}} 
The function \code{parse_utf16_hex} on lines 1-17 takes as its argument
null-terminated pointer \code{s} from which it attempts to read four
characters. These are interpreted as hex digits and converted to an
\code{uint} returned via parameter \code{result}. At first,
\code{s} has no specific bounds annotation, which we can interpret as
\code{count(0)}; this means that \code{s[0]} may be read on line
5. The true branch of the conditional (which extends all the way to
the brace on line 15) is thus typechecked with \code{s} given a
\emph{widened} bound of \code{count(1)}. Likewise, the conditionals on
lines 6-8 each widen it one further; the widened pointer
(\code{s+4}) is returned on success.

The \code{parse} function on lines 18-26 repeatedly invokes
\code{parse_utf16_hex} with its parameter \code{s}, and fills out
array \code{p} whose declared length is the parameter \code{n}. Writes
happen via pointer \code{q}, which is updated using pointer
arithmetic. We specify its bounds as
\code{bounds(p,p+n)} to support this: even as \code{q} changes, 
variables \code{p} and \code{n} do not. Converting from an
\code{array_ptr<uint>} to a \code{ptr<uint>}, done for the call on
line 25, requires proving the array has size at least 1. While this is true
because of the loop condition \code{q < p+n}, which is \code{q}'s
upper bound, the compiler is not smart enough to figure this
out. To convince it, we manually insert a bounds check using
\code{dyn_bounds_cast}.

While bounds checks are \emph{conceptually} inserted on every array
load and store, many of these are eliminated by LLVM\@. For example,
all of the pointer accesses to \code{s} on lines 5-8 are proved safe
at compile-time, so no bounds checks are inserted for
them. \citet{Elliott2018}
reported average run-time overheads of 8.6\% on a pointer-intensive
benchmark suite (49.3\% in one case); \citet{duanrefactoring} measured
no overhead at all on a port of FreeBSD's UDP and IP stacks to \checkedc.

\subsection{Other features}

\checkedc has other features not modeled in this paper. Two in regular
use are \emph{interop types}, which ascribe checked pointer types to
unported legacy code, notably in libraries; and \emph{generic types}
on both functions and \code{struct}s, for type-safe polymorphism. More
details about these can be found in the language specification.

\subsection{Spatial Safety and Backward Compatibility}
\checkedc is backward compatible with legacy C in the sense that all
legacy code will typecheck and compile. However, only code that
appears in \emph{checked regions}, which we call \emph{checked code},
is spatially safe. Checked regions can be designated at the level of
files, functions, or individual code blocks, the first with a
\code{#pragma} and the latter two using the \code{checked}
keyword.\footnote{You can also designate \emph{unchecked} regions
  within checked ones.}  Within checked regions, both
legacy pointers and certain unsafe idioms (e.g., variadic function
calls) are disallowed. The code in Fig.~\ref{fig:checkedc-example}
satisfies these conditions, and will typecheck in a checked region.

How should we think about code that contains both checked and legacy
components? \citet{ruef18checkedc-incr} proved, for a simple
formalization of \checkedc, that \emph{checked code cannot be blamed}:
Any spatial safety violation owes to the execution of unchecked
code. In this paper we extend that result to a richer formalization of
\checkedc.

@@

% There are many existing language development trying to guarantee the C spatial safety property through different approaches. For example, Cyclone \cite{Jim2002} and Cerberus \cite{cerberus} are the fat pointer approaches for guaranteeing the spatial safety and eliminating undefined behaviors in C. Rust \cite{Rust2016}, on the other hand, tries to restrict the possible way of writing programs to guarantee such property statically.
% The problem of these two approaches are effectiveness of executing programs and user practicality of writing programs. The fat-pointer approach completely relies on dynamic checks to guarantee the safety properties but it might pay a huge price on execution efficiency. In our experiment (Sec.\ref{sec:evaluation}), we found that the execution of fat pointer approaches cost the speed a 50\% overhead. 
% On the other hand, Rust uses a static approach by restricting the usage of pointers to guarantee the properties, which results in a user UN-friendly system.  
% The development of \checkedc is an investigation on keeping the balance between execution efficiency and user-friendly systems. We develop a system to utilize static type systems to allow the \checkedc compiler to insert dynamic checks to guarantee the safety property. The efficiency of \checkedc is in the same level of Rust while we maintain a user-friendly system and inherited most of the operation features from C.

@@

% \ignore{
% \section{Null-Terminated Array Pointers}

% Key ideas, formalized:
% \begin{itemize}
% \item Can read one past the end: x[N] can be read, but only written if
%   the value being written is 0.
% \item NT arrays are a refinement of normal arrays: Can convert
%   x:(ntarray N t) to (array N t).
% \item NT arrays can expand their length flow sensitively: Can convert
%   x:(ntarray N t) to (ntarray N+1 t) assuming x[N] != 0.
% \item In general unsafe to convert x:(array N t) to (ntarray N-1 t)
%   even if x[N-1] == 0 because aliases of x could eliminate the NULL
%   terminator. Would need some sort of linearity/alias tracking to
%   support this idea.
% \item Question? What is the syntax of doing a malloc on a NT-array?
%   Should we just do malloc(NT-array l h type) as the malloc of a
%   normal array or should we also include something like int [] nt =
%   “abcd”, where the bound size is not specified?
% \end{itemize}

% See also the Checked C specification
% \ignore{
% https://github.com/microsoft/checkedc/tree/master/spec/bounds\_safety}
% }

\section{Formalization}\label{sec:formal}

% \ignore{
% \begin{figure}
%   \begin{center}
%     \includegraphics[height=6in]{syntax.pdf}
%   \end{center}
%   \caption{\lang: Syntax}
% \end{figure}

>>>>

\section{Overview}\label{sec:overview}

Give an overview of Checked-C and RLBox.

\begin{enumerate}

\item discuss Checked-C and its usage.

\item discuss RLBox.

\item A running example (web-browser with third party libraries).

\end{enumerate}

\section{Key Updates}\label{sec:key-updates}

This section discusses the key merge features between Checked-C and RLBox and the problems we are trying to solve.

\begin{enumerate}

\item Pointing out that the current Checked-C and C formalism is not enough to ensure the security of linking web-browser with third party libraries. Pointing out that even though Checked-C previously has blame theorem, but it still allows program crashing, and our new work can guarantee that every crashes can be detected and handled (non-crashing theorem).

\item Using an example to highlight the overhead of RLBox, and points out the advantage of building RLBox mechanism in a compiler.

\item Discuss function pointer formalism and points out the potential security issues without a proper handling of function pointer set up. Mentioning what happens in RLBox and other tools. Pointing out that this is the first work formalizing function pointers.

\end{enumerate}

% \paragraph*{\textbf{NT-array Pointer Formalization}} NT-array pointers point to NT-arrays whose endings are determined by a null (\code{'\0'}). The length of a NT-array is usually not fixed. The two uses of NT-array pointers in \checkedc are branching and \code{strlen} operations. \code{if (*x) }$e_1$\code{ else }$e_2$ branches to $e_1$/$e_2$ depending one the \code{x}'s data, while \code{strlen(x)} computes the length of the NT-array pointed to by \code{x}.
% \liyi{move to formal and later section.}

% \paragraph*{\textbf{Dependent Functions}} \checkedc allows users to declare dependent functions.  Fig.~\ref{fig:checkedc-example} shows an example usage of dependent functions. The integer argument \code{x} is used to specify the bounds of the two NT-array pointers. It means that when we call the \code{memcpy} function, the upper-bounds of the two NT-array pointers must be no less than \code{x}. If the programmer makes the mistake of writing \lstinline{i <= x} in the for loop, the runtime checks will ensure that the indexing of\lstinline|x| in either \code{a} or \code{b} results in an error.
% Because of the dependent function features, we are able to utilize the fact that the two remaining length of the NT-array pointers are greater than \code{x}, so that the \code{for} is guaranteed to execute properly.

@@

% There are many existing language development trying to guarantee the C spatial safety property through different approaches. For example, Cyclone \cite{Jim2002} and Cerberus \cite{cerberus} are the fat pointer approaches for guaranteeing the spatial safety and eliminating undefined behaviors in C. Rust \cite{Rust2016}, on the other hand, tries to restrict the possible way of writing programs to guarantee such property statically.
% The problem of these two approaches are effectiveness of executing programs and user practicality of writing programs. The fat-pointer approach completely relies on dynamic checks to guarantee the safety properties but it might pay a huge price on execution efficiency. In our experiment (Sec.~\ref{sec:evaluation}), we found that the execution of fat pointer approaches cost the speed a 50\% overhead. 
% On the other hand, Rust uses a static approach by restricting the usage of pointers to guarantee the properties, which results in a user UN-friendly system.  
% The development of \checkedc is an investigation on keeping the balance between execution efficiency and user-friendly systems. We develop a system to utilize static type systems to allow the \checkedc compiler to insert dynamic checks to guarantee the safety property. The efficiency of \checkedc is in the same level of Rust while we maintain a user-friendly system and inherited most of the operation features from C.

@@
